# 1. Randomized Experiments (Week 2)

**New York Scholarship Program**

The following example comes from  Murnane and Willett (2010), chapter 4. The data were originally downloaded from the [UCLA Institute for Digital Research and Education](https://stats.idre.ucla.edu/stata/examples/methods-matter/chapter4/).

The data examine a subset of African-American students from the 1997 New York Scholarship Program, a lottery for school vouchers. 


**Variables**:

  * s_id: id number
  * voucher: received a voucher (1=yes, 0=no)
  * pre_ach: reading achievement score before getting a voucher
  * post_ach: reading achievement score at year 3
  

```{r message=FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(haven)
library(psych)
library(apaTables)
library(gt)
```

```{r message=FALSE, warning=FALSE}
#load data
nysp_vouchers <- read_dta("data/methods_matter/ch4_nyvoucher.dta")
```

```{r message=FALSE, warning=FALSE}
# function to make APA tables
apa <- function(x) {
  gt(x) %>%
  tab_options(
    table.border.bottom.width = 3,
    table.border.bottom.color = "black",
    column_labels.border.top.color = "black",
        column_labels.border.bottom.color = "black",
    table.width = pct(100),
    table.background.color = "white"
  ) %>%
  cols_align(align="center") %>%
  tab_style(
    style = list(
      cell_borders(
        sides = c("top", "bottom"),
        color = "white",
        weight = px(1)
      ),
      cell_text(
        align="center"
      ),
      cell_fill(color = "white", alpha = NULL)
      ),
    locations = cells_body(
      columns = everything(),
      rows = everything()
    )
  )
}

```

## Descriptive Statistics

```{r descriptives, warning=FALSE}
#ADD VARIABLE NAMES TO TABLE
describe(nysp_vouchers) %>%
  as.data.frame() %>%
  mutate(across(is.numeric, round, 3)) %>%
  select(-mad, -min, -max, -range, -trimmed) %>%
  apa()
```

```{r message=FALSE, warning=FALSE}
#combined
rbind(nysp_vouchers %>%
  summarize(voucher = "combined",
            obs = n(),
            mean = mean(post_ach),
            "std. err." = sd(post_ach)/sqrt(n()),
            sd = sd(post_ach)),
  nysp_vouchers %>%
  group_by(voucher) %>%
  summarize(obs = n(),
            mean = mean(post_ach),
            "std. err." = sd(post_ach)/sqrt(n()),
            sd = sd(post_ach))) %>%
  mutate(across(is.numeric, round, 3)) -> desc
  
#differences
diff_data <- data.frame(voucher = "difference",
                      obs = " ",
                      mean = desc$mean[2] - desc$mean[3],
                      `std. err.` = sqrt(
(((nysp_vouchers %>% filter(voucher == 1) %>%
  pull(post_ach) %>% var()) * (nysp_vouchers %>% filter(voucher == 1)  %>% count() %>% pull()-1) +
  (nysp_vouchers %>% filter(voucher == 0) %>%
     pull(post_ach) %>% var()) * (nysp_vouchers %>% filter(voucher == 0)  %>% count() %>% pull()-1)) / (nysp_vouchers %>% count() %>% pull())) *
  ((1/nysp_vouchers %>% filter(voucher == 1)  %>% count() %>% pull())+
  (1/nysp_vouchers %>% filter(voucher == 0)  %>% count() %>% pull()))
),
                      sd = "") %>%
  rename("std. err." = std..err.)
                        
#make a single table
rbind(desc, diff_data) %>% 
  mutate(across(is.numeric, round, 3)) %>%
                      apa()

```


## T-Test

(Strategy 1, Table 4.1, pg. 49)

```{r, mm4_ttest}
mm4_t <- t.test(nysp_vouchers$post_ach ~ nysp_vouchers$voucher, var.equal = T)

broom::tidy(mm4_t) %>%
  mutate(across(is.numeric, round, 3),
         `std. err` = mm4_t[["stderr"]]) %>%
  rename("no voucher" = estimate1,
         "voucher" = estimate2) %>%
  apa()
```


> **Interpretation**
> 
> Students who received an offer of a voucher had significantly higher achievement scores.

## Simple Linear Regression 

(Strategy 2, Table 4.1, pg 49)

```{r, mm4_ols}
#model
mm4_model1 <- lm(post_ach ~ voucher, data=nysp_vouchers)

apa.reg.table(mm4_model1)[[3]] %>%
  apa()
```

### Output for variances 

```{r}
apa.aov.table(mm4_model1)[[3]] %>% 
  apa()
```

> **Interpretation**
> 
> Students who received an offer of a voucher had significantly higher achievement scores (4.9 points higher).

## Multiple Linear Regression 

(Strategy 3, Table 4.1, pg 49)

```{r, mm4_ols2}
mm4_model2 <- lm(post_ach ~ voucher + pre_ach, data = nysp_vouchers)
apa.reg.table(mm4_model2)[[3]] %>% apa()
```

### Output for variances 

```{r}
apa.aov.table(mm4_model2)[[3]] %>% apa()
```

> **Interpretation**
> 
> Students who received an offer of a voucher had significantly higher achievement scores (4.1 points higher) controlling for pre-test scores. The model with the pre_ach covariate is accounts for more variability (is less biased) as indicated by the smaller Mean Square term of 206.58.

*****

**References**

Murnane, R. J., & Willett, J. B. (2010). *Methods matter: Improving causal inference in educational and social science research*. Oxford University Press.