# 7. Propensity Score Matching

## Key Concepts

  * Propensity scores can be used as an attempt to deal with treatment bias in non-experimental data.
  * It is more powerful than simply including many covariates because it attempts to minimize observed differences through predictions from logistic regression.

## Methods Matter, Chapter 12

The following example comes from Murnane and Willett (2010), chapter 12.

**NELS-88 Public vs. Catholic School Math Achievement**

"We investigate the impact on a student’s twelfth-grade mathematics achievement of attending a Catholic (versus a public) high school while attempting to remove bias due to observed differences in base-year family income, parental educational attainment and expectations, and base-year student academic achievement and behavior." This is based on research by Altonji, Elder, and Taber (2005).

The sample contains the 5,671 students from the NELS-88 dataset who were living in families with annual income of less than $75,000 (in 1988 dollars) in the base year of the survey.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(haven)
library(gt)
library(janitor)
library(psych)
library(apaTables)
library(broom)
library(modelsummary)
source("data/apafunction.R")
source("data/apa_msummary.R")
source("data/corstars.R")
```

```{r, message=FALSE, warning=FALSE}
#load and prep data
nels88 <- read_dta("data/methods_matter/ch12_catholic.dta")
```

```{r}
labelled::var_label(nels88) %>%
  unlist() %>%
  matrix(nrow=25, byrow=T) %>% 
  as.data.frame() %>%
  rename("description" = V1) %>%
  add_column(variable = names(nels88), .before=1) %>%
  apa("Key Variables")
  
```

### Descriptive Statistics

```{r message=FALSE, warning=FALSE}
describe(nels88) %>%
  as.data.frame() %>%
  select(n, mean, sd, min, max, skew, kurtosis) %>%
  rownames_to_column("variables") %>%
  slice(3, 6, 17) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  apa("Descriptive statistics for mathematics score (math12), type of high school (catholic), and family income (faminc8)")

nels88 %>%
  mutate(catholic = as_factor(catholic),
         faminc8 = as_factor(faminc8)) %>%
  tabyl(faminc8, catholic) %>%
    adorn_totals(c("row", "col")) %>%
  apa("Catholic school attendance by family income") %>%
  tab_spanner("Attended Catholic school?", vars(no, yes))

nels88 %>%
  tabyl(faminc8, catholic) %>%
  chisq.test()

nels88 %>%
  mutate(faminc8 = as_factor(faminc8),
         catholic = as_factor(catholic)) %>%
  ggplot()+
  coord_flip()+
  geom_point(aes(x=faminc8, y=catholic, color=catholic),
             position = position_jitter())+
  guides(color = guide_legend(reverse=T))+
  labs(title="Distribution of individuals by school type and income") #This is a cool graph and a great visual comparison of the # of individuals in each income bracket
```

#### Generating Variable Categories

The book examples uses generated variables *faminc8*, which uses the actual mid-values of each income range, scaled.

```{r}
nels88 <- nels88 %>%
  mutate(inc8 = dplyr::recode(faminc8,
    "1" = 0,
    "2"=.5,
    "3"=2,
    "4"=4,
    "5"=6.25,
    "6"=8.75,
    "7"=12.5,
    "8"=17.5,
    "9"=22.5,
    "10"=30,
    "11"=42.5,
    "12"=62.5
  ))
```


### Determining the Best Model

Before generating propensity scores, it is important to determine the best model that predicts group membership through an iterative process.

```{r}

# this functions adds chi-square to model summary outputs
glance_custom.glm <- function(x) {
    lrchi2 <- lmtest::lrtest(x)[2,4]
    lrchi2out <- tibble::tibble(`LR Chi-Square:` = lrchi2)
    return(lrchi2out)
}

nels88_modela <- glm(catholic ~ (inc8*math8),
                     data=nels88, family="binomial")

nels88_modelb <- glm(catholic ~ I(inc8^2) + (inc8*math8),
                     data=nels88, family="binomial")

nels88_modelc <- glm(catholic ~ I(inc8^2) + (inc8*math8) + fhowfar + mhowfar + fight8 + nohw8 + disrupt8 + riskdrop8,
                     data=nels88, family="binomial")

msummary(list(
  "Model A" = nels88_modela, 
  "Model B" = nels88_modelb,
  "Model c" = nels88_modelc)) %>%
  apa_msummary("Parameter estimates and approximate p-values for a set of fitted logistic regression models in which attendance at a public or a Catholic high school (CATHOLIC) has been regressed on hypothesized selection predictors (INC8 and MATH8) that describe the base-year annual family income and student mathematics achievement (n = 5,671)")


```
> **Interpretations**
> 
> Model C was determined to be the best based on having smaller model fit statistics AIC/BIC. Also, since Model B is nested within Model C, we can perform a Likelihood-Ratio Chi-squared test between the model deviances (i.e., -2*LL): -2*-1833.541 - -2*-1804.125 = 58.8 which is distibuted as a Chi-squared with degrees of freedom equal the difference in the number of paramters between models (10-4 = 6): $\chi^2$(6)=58.8, p < .001. 

> Why did they include a polynomial? to account for a non-linear relationship between catholic and inc8. See Box-Tidwell (i.e., adding interactions between the continuous variables and its natural log, if it's significant it suggest a non-linearity) and link tests (i.e., this uses the predicted values and square of predicted values to test whether the model is properly specified, the squared term should not be statistically significant) for logistic regression. 

### Examining the Region of Common Support

These graphics show the overlap of the propensity scores of the two groups, suggesting they share a lot of common support on the covariates in the model. Common support indicates the regions of stratification share enough members of the treatment and control groups.


```{r}

# Create Propensity Scores
nels88_propscore <- data.frame(pr_score = predict(nels88_modelc,
                                  type="response")) %>%
  add_column(nels88_modelc$data)

nels88_propscore %>%
  mutate(catholic = ifelse(catholic == 1, "Catholic", "Public")) %>%
  ggplot()+
  geom_histogram(aes(x=pr_score, y=..density..,), color="white", bins=30)+
  geom_density(aes(x=pr_score))+
  facet_wrap(~catholic, ncol=1)

nels88_propscore %>%
  mutate(catholic = ifelse(catholic == 1, "Catholic", "Public")) %>%
  ggplot()+
  geom_histogram(aes(x=pr_score, y=..density.., fill=catholic), color="white", bins=30, position="identity", alpha=.3)+
  geom_density(aes(x=pr_score, color=catholic))#+
  facet_wrap(~catholic, ncol=1) #love this graphic to compare region of common support!
```

### Stratifying on the Propensities

#### Stratification into 5 blocks based on quintiles

```{r}
# making quintiles
nels88_5blocks <- nels88_propscore %>%
  mutate(blocks = ntile(pr_score, 5),
         catholic = factor(catholic, labels = c(
           "Catholic", "Public"
         ))) %>%
  group_by(blocks)
```

#### Checking balance on the propensities
```{r}
do(nels88_5blocks, tidy(t.test(.$pr_score ~ .$catholic,
                 alternative = "greater",
                 paired = FALSE,
                 var.equal = TRUE,
                 conf.level = 0.95
                 ))) %>%
  select(blocks, p.value) %>%
  ungroup() %>%
  apa("Testing balance of blocks based on quntiles")
```

> **Interpretation**
>
> The table above shows t-tests per block between propensity scores and treatment groups. Using a one-sided test where the expectation is a greater score, none of the differences are significant. This suggests balance. 
>
> Note: if I choose "less", I get different results. ?? Yes becuase it is the lower side of the one-sided t-test. You'd be testing whether the expectation is a lower score. 
You may also want to check individual covariate balance, like you do below. 

#### Stratification into 5 blocks based on cut scores

The following stratification comes from p. 320
```{r warning=FALSE}

nels88_5blocks_2 <- nels88_propscore %>%
  mutate(blocks = case_when(
    pr_score < .05 ~ 1,
    pr_score = .05 & pr_score <.1 ~2,
    pr_score =.1 & pr_score <.15 ~ 3,
    pr_score = .15 & pr_score < .2 ~ 4,
    pr_score = .2 & pr_score < .3 ~ 5
  ),
  catholic = factor(catholic, labels = c(
           "Public", "Catholic"
         ))) %>%
  group_by(blocks)

do(nels88_5blocks_2, tidy(t.test(.$pr_score ~ .$catholic,
                 alternative = "greater",
                 paired = FALSE,
                 var.equal = TRUE,
                 conf.level = 0.95
                 ))) %>%
  select(blocks, p.value) %>%
  ungroup() %>%
  apa("Testing balance of blocks based on the book's example")

```

#### Estimate the ATE and ATT of Stratified Propensity Scores

```{r}

# Summarize Block
nels88_5blocks_sum <- nels88_5blocks_2 %>%
  group_by(blocks, catholic) %>%
  summarize(n = n(),
    pr_score = mean(pr_score),
            income_mean=mean(inc8),
            mathmean = mean(math8),
            math_achievment = mean(math12)) %>%
  pivot_wider(names_from=catholic, values_from = c(n, pr_score, income_mean, mathmean, math_achievment)) %>%
  mutate(diff = math_achievment_Catholic - math_achievment_Public,
         n = n_Catholic + n_Public,
         mutate(across(where(is.numeric), round, 3))) %>%
  ungroup()

apa(nels88_5blocks_sum, "Five propensity-score blocks, based on predicted values from final Model C (which contains selected covariates in addition to those included in Model B). Within-block sample statistics include: (a) frequencies, (b) average propensity scores, and (c) average twelfth grade mathematics achievement by type of high school, and their difference (n = 5,671)")

#Weighted Average Treatment Effect
nels88_5blocks_sum %>%
  mutate(weighted = diff * n) %>%
  pull(weighted) %>%
  sum()/sum(nels88_5blocks_sum$n)

#Weighted Average Treatment on the Treated Effect
nels88_5blocks_sum %>%
  mutate(weighted = diff * n_Catholic) %>%
  pull(weighted) %>%
  sum()/sum(nels88_5blocks_sum$n_Catholic)

```



### Using MatchIt

Method from [http://www.practicalpropensityscore.com/matching.html](http://www.practicalpropensityscore.com/matching.html)

```{r message=FALSE, warning=FALSE}
library(MatchIt)

set.seed(2020)

#generate propensity scores using Model C
nels88_ps <- matchit(catholic ~ I(inc8^2) + (inc8*math8) + fhowfar + mhowfar + fight8 + nohw8 + disrupt8 + riskdrop8,
                         data=nels88,
                         method="nearest", #nearest neighbors
                         replace=T, #with replacement
                        ratio=1) #one to one matching
                    
                    #A user may want to consider adding a caliper option to only select cases within a certain range, caliper = .2
       
```


#### Diagnose Covariate Balance

```{r}
nels88_ps_sum <- summary(nels88_ps)

#raw
summary(abs(nels88_ps_sum$sum.matched$`Mean Diff`))

#standardized
summary(abs(summary(nels88_ps, standardize = T)$sum.matched$`Std. Mean Diff`))

table(abs(summary(nels88_ps, standardize = T)$sum.matched$"Std. Mean Diff.") > 0.1) #Table says Max = .05, but this function says there are 11 cases with Std.Mean.Diff > .1

```
> What are we looking for here? Small differences between groups on the covariates. I prefer the standardized differences. These are all farily small (Max = .05). 

#### Estimate Treatment Effects

```{r}
#obtain matched data
matched_data <- match.data(nels88_ps)

# use weights to estimate ATT
library(survey)
design.nels88 <- svydesign(ids=~1, weights=~weights,
                                   data=matched_data)
#estimate the ATT 
model.nels88 <- svyglm(math12~catholic + I(inc8^2) + (inc8*math8) + fhowfar + mhowfar + fight8 + nohw8 + disrupt8 + riskdrop8, design.nels88, family=gaussian())

summary(model.nels88)
```

> The ATT is 1.22 ? Yes, with propensity score matching we can typically only estimate the ATT since we don't typically include the entire sample (see all the unmatched control group cases in the figure below). Weighting by the inverse of the propensity score then running the analysis is one way to get the ATE. You can also get ATE from statification. 


### Another MatchIt Method 

From [Propensity Scores:	A Practical Introduction Using R](https://journals.sfu.ca/jmde/index.php/jmde_1/article/view/431)

#### Visualize Propensity Scores

```{r message=FALSE, warning=FALSE}
plot(nels88_ps, type ="jitter")
```

#### Prepare Matched Data
```{r}
# save matched data
matches <- data.frame(nels88_ps$match.matrix)

# find the  matches
group1 <- match(row.names(matches), row.names(matched_data))
group2 <- match(matches$X1, row.names(matched_data))

# extract outcome values for matches
treatment <- matched_data$math12[group1]
control <- matched_data$math12[group2]

# create matched data set
outcomes <- cbind(matches, treatment, control)

# t-test
t.test(outcomes$treatment, outcomes$control, paired = T) %>% 
  tidy() %>%
  select(estimate:p.value, method, alternative) %>%
  apa("Outcome Analysis")

```
> Is .93 the correct estimate? ATE? ATT? It is very low compared to book. I believe it would be the ATT. While it is lower than what's found in the book, check out the UCLA site, they get estimates this low. One thing we'd want to do is assess the robustness of these results by altering parameters of the matching algorithm (e.g., adding a caliper, altering the values of the caliper .2 to .15, matching with and without replacement, ratio matching 1:5) or by choosing a different matching algorithm (e.g., optimal matching, full matching). These type of tests help demonstrate that results do not depend on the particular matching method chosen.

nels88_matched_data <- match.data(nels88_ps)

### Using PS Match

https://rstudio-pubs-static.s3.amazonaws.com/3473_aec63a6fb6f7437faf2f583233716b92.html

```{r}
# `nonrandom` is not available for R 4.0. You must install like this:
#install.packages(https://cran.r-project.org/src/contrib/Archive/nonrandom/nonrandom_1.42.tar.gz, repos=NULL, type="source")

library(nonrandom)

## Propensity score estimation (pscore object returned)
nels88_psmatch <- pscore(formula  = catholic ~ I(inc8^2) + (inc8*math8) + fhowfar + mhowfar + fight8 + nohw8 + disrupt8 + riskdrop8,
                        data        = nels88,
                        family      = "binomial",
                        name.pscore = "ps")


```

#### Actual Propensity score matching

> Note following code gives error: 

```
Error in ps.match.pscore(object = nels88_psmatch, control.matched.by = matched.by,  : 
  Who was treated? Define argument 'who.treated'.
```
  
```{r eval=FALSE, include=FALSE}
nels88_psmatch_results <- ps.match(object = nels88_psmatch,
                 control.matched.by = matched.by,
                 who.treated = 1, # value 1 is treated
                 name.match.index   = "match.index",
                 ratio              = 1, # default 1:1 match
                 caliper            = "logit",
                 x                  = 0.2,
                 givenTmatchingC    = TRUE,
                 bestmatch.first    = TRUE,
                 setseed            = FALSE,
                 combine.output     = TRUE)
```


## Impact Evaluation, Chapter 8



